---
title: "Brownlows Bets"
author: "Emann Malik"
date: "April 23, 2021"
output: 
  html_document:
    number_sections: true
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(tidyverse)
require(knitr)
```


<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

 At the first day of class there are 35 students in Math Statistics class and professor Brownlow always proposes this bet to his students "I will wager $2 to your bet of $1 that at least two people in today's class have  the same birthday.  He tells me he has made that bet 637 times so far. Where do you think his profits are? Discuss this in a markdown document
 
Report git commit structure:
There are 3 questions and there should be at least 3 separate commits, at least one dealing with each of the 3 questions.
1. What are the fair odds for a bet on matching birthdays in a 35 person class.
2. For what size class is a 2:1 odds bet fair.
3. How much money would Brownlow be expected to end up with if the bets went as described.

# What are the fair odds  


<!--Decide on your methods:  In this section explain the approach you will take to attack the issue --- how will you decide on a likely estimate of professor Brownlow's earnings or losses after 637 bets. Also explain how you construct a 95% confidence interval about your estimate of his winnings !-->
  
```{r}
birthdays = 1:365
saveanswer = vector("numeric")
for(j in 1:1000) {

savebank = 0
for(i in 1:637) {
classsample <- sample(birthdays, 35, replace = TRUE)
d <- duplicated(classsample)

# If we add up 'd' and get a number greater than 0, we have two people who have the same birthday.

if(sum(d) > 0) {
  savebank = savebank + 1
} else {
  savebank = savebank - 4.30
}
}
saveanswer[j] <- savebank
}
```
  
```{r}
hist(saveanswer)
ssanswer <- sort(saveanswer)
print(ssanswer[25])
print(ssanswer[975])
```
  
(The following process was run through for each odds ratio tested.) 
  
1.) The numbers 1 through 365 were created as a vector to represent each day of the year.
  
2.) A simple random sample of 35 with replacement was done to simulate the birthday of each person in a class of 35 students and see whether or not any birthdays were shared.
  
3.) From this sample, any duplicated numbers mean there were people who share the same birthday and were identified and saved. 
  
4.) These duplicate numbers were added together to count the overall amount of students in our simulated class of 35 had matching birthdays. 
  
5.) A loop was created to repeat this sampling process 637 times to see how much money Brownlow made (considering the fact that he gets $1 if he wins and gives $2 if nobody in the class shares a birthday with another student).
  
6.) Another loop was created to run the previous loop 1000 times, meaning 1000 different Brownlows will each have the same bet with the same wager 637 times. 
  
7.) A histogram was created to plot the profit made by each of these 1000 Brownlows after this process was over. The profits followed a normal distribution.
  
8.) The lower (25th Brownlow) and upper (975th Brownlow) 95% confidence interval was calculated to find out how much Brownlow is making (between `r ssanswer[25]` and `r ssanswer[975]`).
  
We have found that, in order to be fair, the odds would have to be exactly or very close to around 4.30:1.
  
# What size of class makes the 2:1 bet fair?  
<!--Explain and display the results of your method of coming up with the estimate of professor Brownlows betting profits!-->
  
```{r}
birthdays = 1:365
saveanswer = vector("numeric")
for(j in 1:1000) {

savebank = 0
for(i in 1:637) {
classsample <- sample(birthdays, 29, replace = TRUE)
d <- duplicated(classsample)

# If we add up 'd' and get a number greater than 0, we have two people who have the same birthday.

if(sum(d) > 0) {
  savebank = savebank + 1
} else {
  savebank = savebank - 2
}
}
saveanswer[j] <- savebank
}
```
  
```{r}
hist(saveanswer)
ssanswer <- sort(saveanswer)
print(ssanswer[25])
print(ssanswer[975])
```
    
(The following process was run through for each class size tested.)
  
1.) The numbers 1 through 365 were created as a vector to represent each day of the year.
  
2.) A simple random sample of size n with replacement was done to simulate the birthday of each person in a class of n students and see whether or not any birthdays were shared.
  
3.) From this sample, any duplicated numbers that mean there are people who share the same birthday were identified and saved. 
  
4.) These duplicate numbers were added together to count the overall amount of students in our simulated classes who had matching birthdays. 
  
5.) A loop was created to repeat this sampling process 637 times to see how much money Brownlow made (considering the fact that he gets $1 if he wins and gives $2 if nobody in the class shares a birthday with another student).
  
6.) Another loop was created to run the previous loop 1000 times, meaning 1000 different Brownlows will each have the same bet with the same wager 637 times. 
  
7.) A histogram was created to plot the profit made by each of these 1000 Brownlows after this process was over. The profits followed a normal distribution.
  
8.) The lower (25th Brownlow) and upper (975th Brownlow) 95% confidence interval was calculated to find out how much Brownlow is making (between `r ssanswer[25]` and `r ssanswer[975]`). 
  
To make the 2:1 bet fair (meaning Brownlow's average profit is 0), the class size would be 28 or 29.  
  
# In the class and payoff as described what is a 95% confidence interval for Brownlow's earnings?
  
<!--Explain and display your method of coming up with a 95% confidence interval of  professor Brownlows winnings (or losings)-->
  
```{r}
birthdays = 1:365
saveanswer = vector("numeric")
for(j in 1:1000) {

savebank = 0
for(i in 1:637) {
classsample <- sample(birthdays, 35, replace = TRUE)
d <- duplicated(classsample)

# If we add up 'd' and get a number greater than 0, we have two people who have the same birthday.

if(sum(d) > 0) {
  savebank = savebank + 1
} else {
  savebank = savebank - 2
}
}
saveanswer[j] <- savebank
}
```
  
```{r}
hist(saveanswer)
ssanswer <- sort(saveanswer)
print(ssanswer[25])
print(ssanswer[975])
```
  
1.) The numbers 1 through 365 were created as a vector to represent each day of the year.
  
2.) A simple random sample of 35 with replacement was done to simulate the birthday of each person in a class of 35 students and see whether or not any birthdays were shared.
  
3.) From this sample, any duplicated numbers that mean there are people who share the same birthday were identified and saved. 
  
4.) These duplicate numbers were added together to count the overall amount of students in our simulated class of 35 had matching birthdays. 
  
5.) A loop was created to repeat this sampling process 637 times to see how much money Brownlow made (considering the fact that he gets $1 if he wins and gives $2 if nobody in the class shares a birthday with another student).
  
6.) Another loop was created to run the previous loop 1000 times, meaning 1000 different Brownlows will each have the same bet with the same wager 637 times. 
  
7.) A histogram was created to plot the profit made by each of these 1000 Brownlows after this process was over. The profits followed a normal distribution.
  
8.) The lower (25th Brownlow) and upper (975th Brownlow) 95% confidence interval was calculated to find out how much Brownlow is making. 
  
The 95% confidence interval for the profit Brownlow is making from this bet is between `r ssanswer[25]` and `r ssanswer[975]`. 
  
# Discussion and Conclusion

<!--Discussion and conclusion here.  If professor Brownlow is consistantly making or losing money how could he change his payout so that his profits center on zero?  Also discuss what size his class should be in order to have his current scheme balance at zero profit.!-->
  
In conclusion, we have found that:      
  
1.) In order to be fair, the odds would have to be exactly or very close to around 4.30:1.  
  
2.) To make the 2:1 bet fair (meaning Brownlow's average profit is 0), the class size would be 28 or 29.     
  
3.) The 95% confidence interval for the profit Brownlow is making from this bet is between `r ssanswer[25]` and `r ssanswer[975]`.    
  